---
title: "Using TimeSeries.OBeu with OpenCPU"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This document describes the use of the functions implemented in *TimeSeries.OBeu* package in *OpenCPU* environment, after *[installing OpenCPU](https://www.opencpu.org/download.html)* and *TimeSeries.OBeu* package on your OpenCPU server address(/ocpu/test).

# How to use functions

- Go to: yourserver/ocpu/test

- Type to the endpoint:

```{r, eval=FALSE, include=TRUE}
 ../library/ {name of the library} /R/ {function}
```

- If you want to see the function parameters you should: 

    - Select Method: **`Get`**
    
- To run a function you should:

    - Select Method: **`Post`**

- Then you can click on **Ajax Request**.

# OpenCPU and TimeSeries.OBeu

## ts.analysis

In this example we will use `ts.analysis` function that returns in a single call a json string or a list with the following components:

+-----------------------------+---------------------------+-------------------------------------------------------------+
| Component                   | Output                    | Description                                                 |
+=============================+===========================+=============================================================+
| acf.parameters              | - acf                     | - ACF values of the input time series                       |
|                             | - acf.lag                 | - Lags at which the acf is estimated                        |
|                             | - confidence.interval.up  | - Upper limit of the confidence interval                    |
|                             | - confidence.interval.low | - Lower limit of the confidence interval                    |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| pacf.parameters             | - pacf                    | - PACF values of the input time series                      |
|                             | - pacf.lag                | - Lags at which the pacf is estimated                       |
|                             | - confidence.interval.up  | - Upper limit of the confidence interval                    |
|                             | - confidence.interval.low | - Lower limit of the confidence interval                    |
+-----------------------------+---------------------------+-------------------------------------------------------------+
|  acf.residuals.parameters   | - acf.res                 | - ACF values of the model residuals                         |
|                             | - acf.res.lag             | - Lags at which the acf is estimated of the model residuals |
|                             | - confidence.interval.up  | - Upper limit of the confidence interval                    |
|                             | - confidence.interval.low | - Lower limit of the confidence interval                    |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| pacf.residuals.parameters   | - pacf.res                | - Pacf values of the model residuals                        |
|                             | - pacf.res.lag            | - Lags at which the pacf is estimated of the model residuals|
|                             | - confidence.interval.up  | - Upper limit of confidence interval                        |
|                             | - confidence.interval.low | - Lower limit of confidence interval                        |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| stl.plot                    | - trend                   | - Trend component                                           |
|                             | - trend.ci.up             | - Up limit for trend component                              |
|                             | - trend.ci.low            | - Low limit for trend component                             |
|                             | - seasonal                | - Seasonal component                                        |
|                             | - remainder               | - Remainder component                                       |
|                             | - time                    | - Time of the series was sampled                            |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| stl.general                 | - stl.degree              | - Degree of fit                                             |
|                             | - degfr                   | - Effective degrees of freedom                              |
|                             | - degfr.fitted            | - Fitted degrees of freedom                                 |
|                             | - fitted                  | - Model's fitted values                                     |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| residuals                   | - residuals               | - Residuals of the model                                    |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| compare                     | - arima.order             | - Arima order                                               |
|                             | - arima.coef              | - AR, MA and regression coefficients                        |
|                             | - arima.coef.se           | - Standard error of the coefficients                        |
|                             | - covariance.coef         | - Variance of the coefficients                              |
|                             | - resid.variance          | - Residuals variance                                        |
|                             | - not.used.obs            | - Number of not used observations                           |
|                             | - used.obs                | - Used observations                                         |
|                             | - loglik                  | - Maximized log-likelihood,                                 |
|                             | - aic                     | - AIC value                                                 |
|                             | - bic                     | - BIC value                                                 |
|                             | - gcv                     | - Generalized cross-validation statistic                    |
|                             | - aicc                    | - Second-order AIC                                          |
+-----------------------------+---------------------------+-------------------------------------------------------------+
| forecasts                   | - ts.model                | - A string indicating the arima orders                      |
|                             | - data_year               | - Time of time series data                                  |
|                             | - data                    | - Time series values                                        |
|                             | - predict_time            | - Time of the predicted values                              |
|                             | - predict_values          | - Predicted values                                          |
|                             | - up80                    | - Upper 80% confidence limit                                |
|                             | - low80                   | - Lower 80% confidence limit                                |
|                             | - up95                    | - Upper 95% confidence limit                                |
|                             | - low95                   | - Lower 95% confidence limit                                |
+-----------------------------+---------------------------+-------------------------------------------------------------+

Table: *`ts.analysis`* components

## Select library and function

1. Go to: yourserver/ocpu/test

2. Copy and paste the following function to the endpoint 
```{r, eval=FALSE, include=TRUE}
../library/TimeSeries.OBeu/R/ts.analysis
# library/ {name of the library} /R/ {function}
```

3. **Select Method**: **`Post`**

## Adding parameters parameters

Click **add parameters** every time you want to add a new parameters and values.

4. Define the input data:

    - **Param Name**: **`tsdata`**
    - **Param Value**: e.g. **`Athens_executed_ts`** 

5. Define the *prediction steps* parameter:

    - **Param Name**: **`prediction.steps`**
    - **Param Value**: **`2`**

You add likewise *x.order* parameter to fit a specific arima order, see TimeSeries.OBeu *reference manual* for further details.

6. Ready! Click on **Ajax request**!

## Results

7. copy the **/ocpu/tmp/{this_id_number}/R/.val** (second on the right panel)

8. finally, paste **`yourserver/ocpu/tmp/{this_id_number}/R/.val`** on a new tab.


# Further Details

+ [HTTP in OpenCPU](https://www.opencpu.org/api.html)
+ [OpenCPU Help](https://www.opencpu.org/help.html)
+ [OpenCPU JavaScript Client](https://www.opencpu.org/jslib.html)
+ [OpenCPU on CRAN](https://cran.r-project.org/package=opencpu)

# Github
+ [OpenCPU package *development version*](https://github.com/opencpu/opencpu)
+ [TimeSeries.OBeu package *development version*](https://github.com/okgreece/TimeSeries.OBeu)
